---
title: "data_wrangling_i"
author: "Courtney Chan"
date: "September 18, 2018"
output: html_document
---

#Lecture
What is a tibble
Dataframe - holds lists
  stores information of multiple variable types - numbers and characters

With tibble package - layers on top of R, with a similar data structure
  Tibble is a version of a dataframe gives a shortened summary

multiple waysto import data
how to code missing values = 99 or . etc

Readable code for data cleaning -ither ppl can understand it

library(tidyverse)

# Paths

paths to files in your computer versus how R knows what the file is
GitWD - working directory the folder that contains the R project

Different methods of giving your computer a path
Absolute path: full address of the file
computer's home directory

Ex. "/Users/jeffgoldsmith/Dropbox/Work/Teaching/P8105/Fall_2018/p8105.github.io/data/FAS_litters.csv"

From a starting point

Inserting this absolute path in your document - this will change if you give this file to someone else

Versus Relative path: give directions from where you are right now

Relative paths are better

"./data/FAS_litters.csv"

Would need the directory subfolder "data" that contains the file you need - can send the whole R project which should contain the subfolder you need?

We know already we are starting from this R project - then the list is relative to this project - can send to someone else and ti would still work

the "." designates wherever you are, then go to the working directory/the specific data file

set(wd) is a way of setting the absolute path -so... don't do it

"~" refers to home directory - will be used if you're using absolute paths


put the Rmd file within the sub-directory folder
can use the here::here when you dont' have your Rmd file not in your sub-directory - but this can be more complicated

##import FAS csv files

can insert the absolute path or the relaitve path in file =

```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
```

getwd() tells you what working directory you're in

"C:/Users/chanc/Documents/Fall 2019/DataScience I/Lectures/Lecture 4 9-18-18/data_wrangling_i"

R doesn't do well with column names with spaces - can use 
litters_data = janitor::clean_names(litters_data)

janitor is the package and the clean names is the function - so that it won't conflict with other packages that have functions of the same name - so don't have to load the whole janitor package when you have other packages already loaded

Janitor::clean_names gets rid of the original names and overwrites with cleaner names, such as lower snake case

```{r}
litters_data = read_csv(file = "./data/FAS_pups.csv")
litters_data = read_csv(file = "/Users/chanc/Documents/Fall 2019/DataScience I/Lectures/Lecture 4 9-18-18/data_wrangling_i/data/FAS_pups.csv")
getwd()
```

Default is to print the first 10 and last 10 rows - should check each time you import so you are using the right dataset

head(litters_Data)
tail(litters_data)

skimr::skim you can see mini histograms for each variable


View() function allows you to view dataset - pulls up all the columns and rows. Make sure your data is being changed accordingly

Under the read_csv, there are additional arguments
skip = 0 whether you should skip some columns that are empty
na: 
col_names = TRUE, that the first row is variable names. If false, will just create its own column names using X

col_types, based on values in the columns, will determine what the datatypes are by itself
assume that column is numeric etc.
occasionally it will be wrong

```{r}
head(litters_data)
tail(litters_data)
skimr::skim(litters_data)
```


col_types forces the column types to be what you designate
col_types = "ciiiii""
character and integers
say making litter number is designated as character, minimum number of times it was repeated in the set, was 3, max times a litter number was repeated was 15 times

forcing something to be another variable type might be wrong - shows up as NA

can't use csv if it is a xlsx
don't want to save a xlsx spreadsheet into a csv
rather should use readxl package
reads xl files especially

mlb_data= readxl::read_excel(path "")

can tell it to read an excel range
mlb_subset = range = "A1:E17"
gives smaller collection of data given the cell range specified


how to import SAS dataset
pulse_Data = haven::read_sas("")
haven is the package and read_sas is the function

##compare with base R

pups_readr
pups_baser

readr produces tibble
while baser users just underlying R dataframe

printing baser, will print out everything
while readr won't

pups_baser$S can extract this one variable you are designating, here you say "S"
will extract the entire column

while read$S is not a valid function
